```{r echo=FALSE, results="hide"}
library(knitr)
library(plyr)
opts_chunk$set(tidy=FALSE)
source('../R/functions-analyses.R')
source('../R/functions-figures.R')
```

# Examples from lesson "Intro to R" - presented by Diego Barneche  

## Vector types

There are 5 different types of vector in R.  
```{r}
x <- TRUE       # x becomes TRUE
x <- 1          # x now has a value of 1 (overwritten the line above)
x <- c(1, TRUE) # x now contains two values of 1, because a vector always stores the same mode of elements
x
x <- c(1, TRUE, '1') #now all values are characters, because there is a underlying hierarchy. For instance, if there is one element of mode 'character' in your vector, then all elements will be treated as so
x
```

## Lists  
If you want to store the different modes of elements in one single object, you use lists.  
```{r}
x <- list(1, TRUE, "1")
x
```

## Querrying  
You can ask yourself about the nature of an unknown vector - suppose you did not know what x is:  
```{r}
storage.mode(x)
typeof(x)
class(x)
```

# Examples from lesson "Writing functions" - presented by Dan Warren  

## Playing with the basics  
When you write your function, you define it based on *arguments*  
```{r}
f <- function(something) {
    something
}
f(100)
```
The function `f()` is what we call the identity function, i.e., the function returns the exact input. In this particular case, the argument `something` took the value 100. You could instead define an object before, and then calling the function on that object, which should do exactly the same thing.  
```{r}
x  <-  100 #x now is a numeric vector of one element of value 100
f(x)
```
It returned exactly the same thing. In this case, `something` took the value of x, which was 100.  

## Downloading the gapminder data
We have to download a .csv file that we will use for the rest of our course. Right-click this [link][id] and save it to some new directory. Let's play with it now. 
```{r}
data  <-  read.csv('../data/gapminder-FiveYearData.csv', header=TRUE, stringsAsFactors=FALSE)
head(data) #try the function tail() instead, what happened?

#subset the data for year 1982 only
data.1982  <-  data[data$year == 1982, ]

#let's create a function to compute the arithmetic average of a vector
average  <-  function(x) {
    sum(x) / length(x)
}

average(data.1982$gdpPercap)
#same as 
sum(data.1982$gdpPercap) / length(data.1982$gdpPercap)
```
Notice that repeating things may be problematic; first because it increases your chances of typos; second, if for some reason you change the name of your columns in `dat`, you will have to manually change all these names, which will be quite annoying.  

# Examples from lessons "Control Flow & Repeating things" - Writing plot functions and looping - presented by Dan Warren & Diego Barneche  

## Making a plot  
```{r echo=FALSE, fig.width=10, fig.height=8, message=FALSE, warning=FALSE}
allNiceBoxPlots(data)
```

[id]: https://raw.github.com/dbarneche/2014-07-14-Dalhousie/gh-pages/data/lessons/10-functions/gapminder-FiveYearData.csv